Reproducible methods for modeling combined public transport and cycling trips and associated benefits: \ evidence from the biclaR tool

### 2024-10-06 paper/PaperCEUS/PaperCEUS_rev2_word.Rmd
tiny typos

```{r eval=FALSE, echo=FALSE}
# setwd("paper/PaperCEUS/") # knit directory: Document
system("latexdiff paper/PaperCEUS/PaperCEUS_rev1.tex paper/PaperCEUS/PaperCEUS_rev2.tex > paper/PaperCEUS/PaperCEUS_diff2.tex")
#and change in tex the blue for another color (green? 0,0.5,0) and remove \uwave
# \definecolor{green}{rgb}{0,0.5,0}
# {\protect\color{green}{#1}}}
#and compile PDF!
system("pdflatex paper/PaperCEUS/PaperCEUS_diff2.tex")
```

@@

Using the [`odjitter` R package](https://github.com/dabreegster/odjitter), we disaggregated the OD data into desire lines reprenting a maximum of 100 trips each.
Figure \ref{fig:jitter} illustrates the contrast between trip representation through the traditional method, which connects a single desire line between each district, and the presentation achieved through the randomization and disaggregation of trips between districts, specifically for the Lisbon metropolitan area.
As shown, the city of Lisbon (in the centre) is the main attractor of trips, with a high number of trips to and from the other municipalities.

@@

To estimate the car emissions, we used the EMEP/EEA's COPERT software v5 methods and reference values [@COPERT] for a Tier 3 detail level. 
We used a family-size vehicle, EURO standard, and gasoline or diesel fuel.
All trips were considered to be made under urban conditions and at an average speed of 15 km/h during rush hour periods.
Since the average distance traveled per trip influences the overconsumption and emissions from cold-start engine operation, we estimated energy and emission factors for different ranges of trips at 500-meter intervals.  
An equation was then used to calculate emission factors for the two types of fuel, for each type of pollutant, whose explanatory variables are driving speed ($speed$, in km/h) and average trip distance ($l_{trip}$, in km/trip). Thus, the emission factors ($EF_{fuel,l_{trip},speed}$, in g/km) can be calculated using equation \ref{eq:fatoremissaoauto}.

@@

knitr::kable(summary1,
      digits = 0,
      format.args = list(big.mark = " "),
      col.names = c("Target", "Routing", "Total trips",
                    "Baseline Cycling + PT", "Potencial Cycling + PT"),
      caption = "\\label{summary1}Summary of the cycling potencial of the intermodality scenario. Values in 'trips/day'.",
      format="latex",
      booktabs = TRUE) |> 
  kableExtra::column_spec(column = c(4:5), width = "7em")

@@

knitr::kable(summary1b,
      digits = c(0,0,1,0,0,0),
      format.args = list(big.mark = " "),
      col.names = c("Target", "Routing",
                    "Avoided Mortality (deaths/yr)", "Social benefits (k€/yr)", 
                    "Avoided CO2eq (ton/yr)", "Environmental benefits (k€/yr)"),
      caption = "\\label{summary1b}Summary of the cycling potencial of intermodality scenario and its socio-environmental benefits for the cycling legs.",
      format="latex",
      booktabs = TRUE) |> 
  kableExtra::column_spec(column = c(3:6), width = "6em")

@@

knitr::kable(summary22,
      digits = c(0,0,0,0,1,0,1,0),
      format.args = list(big.mark = " "),
      col.names = c("Target", "Routing", "CO2eq",
                    "CO", "PM10", "NOx", "VOC", "Value (k€)"),
      caption = "\\label{summary22}Summary of the avoided emmissions (ton/year) and corresponding monetization (thousand €) by replacing car trips with PT, in the second leg.",
      format="latex",
      booktabs = TRUE) |> 
  kableExtra::column_spec(column = c(3:7), width = "3.5em")
```

@@

knitr::kable(summaryall,
      digits = 0,
      format.args = list(big.mark = " "),
      col.names = c("Target", "Routing", "Avoided CO2eq (tons)",
                    "SE Benefits 1yr (k€)", "SE Benefits 10yrs (k€)"),
      caption = "\\label{summaryall}Summary of the avoided CO2eq emmissions (ton/year) and the estimated social and environmental benefits (monetized in thousand €) by replacing car trips with cycling in combination with PT.",
      format="latex",
      booktabs = TRUE) |> 
  kableExtra::column_spec(column = c(3:5), width = "8em")

@@

The environmental impacts represent less than 2% of the socio-environmental benefits (in value) from replacing car trips to bicycle in first-and-last legs. 
For the PT segment, we did not estimate the social impacts from substituting car trips. 
One of the main socio-environmetal benefits, valued after monetization, comes from the increase in physical activity [@Felix2023ES].
Although there are also social benefits form shifting car trips to PT, its health benefits would not be as high as shifting to cycling. 
The literature shows that the Metabolic Equivalent Tasks (MET) for "riding in a bus or a train" is 1.3 plus the "walking for transportation" as 3.5, while "driving a car" is 2.5 [@MET2011].
The difference between these activities - shifting from car to PT - is not very obvious when compared to shifting from car to cycling, whose MET is about 6.8.
Nevertheless, future works should also encompass the estimation of the social impacts for the PT leg of the journey, shifting from car.
 <!--Overall there was very little evidence available on mode switch to public transport. -->

@@

The findings indicate that cycling combined with public transport has the potential to account for 10% of total trips. Transitioning to cycling for the first and last segments could lead to a decrease in annual CO~2~eq emissions by a substantial margin, ranging from 3,000 to 7,500 tons per day, depending on the cycling target and routing profile. Moreover, for the second trip segment, the shift from car to public transport contributes to avoiding up to 20,500 tons of CO~2~eq emissions annually. These changes are estimated to yield socio-environmental benefits totaling €125 million to €325 million over a decade.
The quantification of such benefits can support policy-makers in prioritizing interventions to reduce the reliance on private motorized modes of transportation.

>>>>

```{r eval=FALSE, echo=FALSE}
# setwd("paper/PaperCEUS/") # knit directory: Document
system("latexdiff paper/PaperCEUS/PaperCEUS_rev1.tex paper/PaperCEUS/PaperCEUS_rev2_word.tex > paper/PaperCEUS/PaperCEUS_diff2.tex")
#and change in tex the blue for another color (green? 0,0.5,0) and remove \uwave
# \definecolor{green}{rgb}{0,0.5,0}
# {\protect\color{green}{#1}}}
#and compile PDF!
system("pdflatex paper/PaperCEUS/PaperCEUS_diff2.tex")
```

@@

Using the [`odjitter` R package](https://github.com/dabreegster/odjitter), we disaggregated the OD data into desire lines representing a maximum of 100 trips each.
Figure \ref{fig:jitter} illustrates the contrast between trip representation through the traditional method, which connects a single desire line between each district, and the presentation achieved through the randomization and disaggregation of trips between districts, specifically for the Lisbon metropolitan area.
As shown, the city of Lisbon (in the centre) is the main attractor of trips, with a high number of trips to and from the other municipalities.

@@

To estimate the car emissions, we used the EMEP/EEA's COPERT software v5 methods and reference values [@COPERT] for a Tier 3 detail level. 
We used a family-size vehicle, EURO standard, and gasoline or diesel fuel.
All trips were considered to be made under urban conditions and at an average speed of 15 km/h during rush hour periods.
Since the average distance travelled per trip influences the overconsumption and emissions from cold-start engine operation, we estimated energy and emission factors for different ranges of trips at 500-meter intervals.  
An equation was then used to calculate emission factors for the two types of fuel, for each type of pollutant, whose explanatory variables are driving speed ($speed$, in km/h) and average trip distance ($l_{trip}$, in km/trip). Thus, the emission factors ($EF_{fuel,l_{trip},speed}$, in g/km) can be calculated using equation \ref{eq:fatoremissaoauto}.

@@

knitr::kable(summary1,
      digits = 0,
      format.args = list(big.mark = " "),
      col.names = c("Target", "Routing", "Total trips",
                    "Baseline Cycling + PT", "Potential Cycling + PT"),
      caption = "\\label{summary1}Summary of the cycling potential of the intermodality scenario. Values in 'trips/day'.",
      format="latex",
      booktabs = TRUE) |> 
  kableExtra::column_spec(column = c(4:5), width = "7em")

@@

knitr::kable(summary1b,
      digits = c(0,0,1,0,0,0),
      format.args = list(big.mark = " "),
      col.names = c("Target", "Routing",
                    "Avoided Mortality (deaths/yr)", "Social benefits (k€/yr)", 
                    "Avoided CO2eq (ton/yr)", "Environmental benefits (k€/yr)"),
      caption = "\\label{summary1b}Summary of the cycling potential of intermodality scenario and its socio-environmental benefits for the cycling legs.",
      format="latex",
      booktabs = TRUE) |> 
  kableExtra::column_spec(column = c(3:6), width = "6em")

@@

knitr::kable(summary22,
      digits = c(0,0,0,0,1,0,1,0),
      format.args = list(big.mark = " "),
      col.names = c("Target", "Routing", "CO2eq",
                    "CO", "PM10", "NOx", "VOC", "Value (k€)"),
      caption = "\\label{summary22}Summary of the avoided emissions (ton/year) and corresponding monetization (thousand €) by replacing car trips with PT, in the second leg.",
      format="latex",
      booktabs = TRUE) |> 
  kableExtra::column_spec(column = c(3:7), width = "3.5em")
```

@@

knitr::kable(summaryall,
      digits = 0,
      format.args = list(big.mark = " "),
      col.names = c("Target", "Routing", "Avoided CO2eq (tons)",
                    "SE Benefits 1yr (k€)", "SE Benefits 10yrs (k€)"),
      caption = "\\label{summaryall}Summary of the avoided CO2eq emissions (ton/year) and the estimated social and environmental benefits (monetized in thousand €) by replacing car trips with cycling in combination with PT.",
      format="latex",
      booktabs = TRUE) |> 
  kableExtra::column_spec(column = c(3:5), width = "8em")

@@

The environmental impacts represent less than 2% of the socio-environmental benefits (in value) from replacing car trips to bicycle in first-and-last legs. 
For the PT segment, we did not estimate the social impacts from substituting car trips. 
One of the main socio-environmental benefits, valued after monetization, comes from the increase in physical activity [@Felix2023ES].
Although there are also social benefits form shifting car trips to PT, its health benefits would not be as high as shifting to cycling. 
The literature shows that the Metabolic Equivalent Tasks (MET) for "riding in a bus or a train" is 1.3 plus the "walking for transportation" as 3.5, while "driving a car" is 2.5 [@MET2011].
The difference between these activities - shifting from car to PT - is not very obvious when compared to shifting from car to cycling, whose MET is about 6.8.
Nevertheless, future works should also encompass the estimation of the social impacts for the PT leg of the journey, shifting from car.
 <!--Overall there was very little evidence available on mode switch to public transport. -->

@@

The findings indicate that cycling combined with public transport has the potential to account for 10% of total trips. Transitioning to cycling for the first and last segments could lead to a decrease in annual CO~2~eq emissions by a substantial margin, ranging from 3,000 to 7,500 tons per day, depending on the cycling target and routing profile. Moreover, for the second trip segment, the shift from car to public transport contributes to avoiding up to 20,500 tons of CO~2~eq emissions annually. These changes are estimated to yield socio-environmental benefits totalling €125 million to €325 million over a decade.
The quantification of such benefits can support policy-makers in prioritizing interventions to reduce the reliance on private motorized modes of transportation.
